---
title: "Untitled"
runtime: shiny
output: 
 flexdashboard::flex_dashboard:
  orientation: columns
  vertical_layout: scroll
  source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(shinyjs)
library(shinyjs)
library(lobstr)
library(htmltools)
library(readr)
library(magrittr)
library(DT)
library(ggplot2)
library(plotly)
library(quantmod)
library(forecast)
library(tseries)
library(fGarch)
library(rugarch)
```

# PAGE1 {data-icon="fa-table"}

## Column_1 {.sidebar}

```{r}
useShinyjs(rmd = TRUE)

fileInput(inputId="Upload_1", label="Choose the file", multiple = FALSE, accept = ".csv")

textInput(inputId="Upload_2", label="Symbol", value="")
dateRangeInput("Dates_set", label = h3("Date range"), start = Sys.Date()-800, end = Sys.Date())

actionButton("reset", label = "Reset", width="80px")
actionButton("submit", label = "Submit", class="btn btn-primary")
```


```{r}
observeEvent(input$submit11,{
  shinyjs::reset("Dates_plot")
  value$Date_min<-NULL
  value$Date_max<-NULL
})
```



```{r}
#evaluation featres of isTruthy() - https://cran.r-project.org/web/packages/shiny/shiny.pdf
observeEvent(input$submit,{
  
  if (fileCheck() && input$Upload_2=="") {
    
      Import1<-read_csv(input$Upload_1$datapath)
    if(ncol(Import1)==1){
      Import1<-read_csv2(input$Upload_1$datapath)
    }
      value$data<-as.data.frame(Import1)
  }
  else if (isFALSE(fileCheck()) && isTruthy(input$Upload_2)==TRUE){
    
    tryCatch({
      if(is.na(input$Dates_set[1]) || is.na(input$Dates_set[2])){
        Import2<-getSymbols(input$Upload_2, src="yahoo", auto.assign = FALSE)
      }
      else{
        Import2<-getSymbols(input$Upload_2, from=input$Dates_set[1], to=input$Dates_set[2]+1,  src="yahoo", auto.assign = FALSE)
      }
      
      value$data<-data.frame(Date=index(Import2), coredata(Import2)) },
      
    error = function(c) {runjs('alert("Error: symbol does not exist.");')},
    finally={gc()}
    )
  }
  else { 
    erase()
  }
})
```

```{r}
observeEvent(input$reset,{erase()})
```

```{r}
erase<-function() {
    reset("Upload_1")
    reset("Upload_2")
    value$addr_old <- value$addr_new
    runjs('alert("Controls reset!");')
}
```



```{r} 
#Data storage
value <- reactiveValues(
  data = NULL,
  addr_old=NULL,
  addr_new=NULL,
  DateColInd=NULL,
  Date_min=NULL,
  Date_max=NULL,
  choise=NULL,
  series=NULL
)
```

```{r}
#block of code, that activates once. Needed for fileInput control reset
observeEvent(handlerExpr=NULL,{
  value$addr_new<-obj_addr(input$Upload_1)
  value$addr_old<-obj_addr(input$Upload_1)}, once = TRUE)
```

```{r}
#Here we check, whether fileInput has been updated.
fileCheck<-eventReactive(c(input$Upload_1, value$addr_old),{
  
  value$addr_new <- obj_addr(input$Upload_1)
  
  if(value$addr_new == value$addr_old) {FALSE}
  else {
    value$addr_old == value$addr_new
    TRUE
  }
})
```








```{r}
#Initialization block
observeEvent(value$data,{
  req(is.data.frame(value$data)==TRUE)
  
  value$data<-na.omit(value$data) #modify a bit
  
  is.date <- function(x) inherits(x, 'Date')
  index <- sapply(value$data, is.date) %>% which
  
  if(length(index) == 1) {
    value$Date_min <- value$data[1,index]
    value$Date_max <- value$data[nrow(value$data),index]
    value$DateColInd<-index
  }
  else {
    
    is.convertible.to.date <- function(x) !anyNA(as.Date(as.character(x), tz = 'UTC', format = '%Y-%m-%d'))
    index <- sapply(value$data, is.convertible.to.date) %>% which
    
    if(length(index) != 1) {
      
      ### make an alert
    }
    else {
      value$Date_min <- value$data[1,index]
      value$Date_max <- value$data[nrow(value$data),index]
      value$DateColInd<-index
    }
  }
  

  updateDateRangeInput(session, "Dates_plot",
    start= value$Date_min,
    end= value$Date_max
   )

  
  value$choise <- as.list(colnames(value$data)[unlist(lapply(value$data, is.numeric), use.names = FALSE)])
                          
  updateSelectizeInput(session, "select", choices = value$choise, selected = NULL, options = list(), server = FALSE)
  
  updateSelectizeInput(session, "finalSelect", choices = value$choise, selected = NULL, options = list(), server = FALSE)

})

```


```{r}
observe({
  req(is.null(value$data)==FALSE) #TryCatch - не работает, необходио дописать! 
  
    if (input$Finaldates[1]<input$Finaldates[2]){
      colIndex <- colnames(value$data) %>% grepl(input$finalSelect,.) %>% which
      value$series <- value$data[(value$data[,value$DateColInd]>=input$Finaldates[1] & value$data[,value$DateColInd]<=input$Finaldates[2]), colIndex] 
    } else {
      colIndex <- colnames(value$data) %>% grepl(input$finalSelect,.) %>% which
      value$series <- value$data[ , colIndex]    
    }
})  
```






## Column_2

### Time series plot
```{r}
fillCol(flex=c(8.5,1.5),
  plotlyOutput("MainPlot", height="100%"),
  
  fillRow(
    dateRangeInput("Dates_plot", label = "Date range"),
    selectizeInput("select", label = "Select box", 
    choices = NULL,
    selected = 1, options=list(dropdownParent="body"))
  )
)
```


```{r}
output$MainPlot <-renderPlotly({
  
  req(is.data.frame(value$data), isTruthy(value$choise))
  p<-ggplot(value$data)+aes_string(x=value$data[,value$DateColInd], y= input$select) + coord_cartesian(xlim = c(input$Dates_plot[1],input$Dates_plot[2])) + geom_line()
  plotly::ggplotly(p)
  
}) 
```



### Data transformation panel  

```{r}
fillRow(
  selectizeInput("finalSelect", label = h3("Series to analyze"), 
      choices = NULL,
      selected = 1, options=list(dropdownParent="body")),
  
  dateRangeInput("Finaldates", label = h3("Range")), height = "100px"
)

checkboxGroupInput("checkGroup", label = h3("Checkbox group"), 
    choices = list("Outlier" = 1, "Round" = 2, "Log" = 3),
    selected = 1)

```


```{r}
observe({
   updateDateRangeInput(session, "Finaldates",
    start= value$Date_min,
    end= value$Date_max
 )
})

```



### Datatable
```{r}
output$DataTable<- renderDataTable(

  datatable(value$data, filter="top") #%>% 
    #formatRound(colnames(value$data)[unlist(lapply(value$data, is.numeric), use.names = FALSE)], 2)
)
```

```{r}
dataTableOutput('DataTable')
```



```{r}
output$a111<- renderPrint({mData$Tseries})
verbatimTextOutput("a111")
```



# ARIMA/ARCH/GARCH

```{r}
mData <- reactiveValues(
  Tseries=NULL,
  arima=NULL,
  p=NULL,
  q=NULL,
  d=NULL,
  adf=NULL,
  diffSeries=NULL,
  LMtestTable=NULL,
  ArchModels=NULL,
  crARCH=NULL,
  GarchModels=NULL,
  crGARCH=NULL,
  TGarchModels=NULL,
  crTGARCH=NULL,
  MGarchModels=NULL,
  crMGARCH=NULL
)
```

```{r}
actionButton("button1", "Analyze")
```

### Foreword

In financial time series, it is often that the series is transformed by logging and then the differencing is performed.People often look at the returns of the stock rather than the its prices. Differences of log prices represent the returns and are similar to percentage changes of stock prices.


### Estimated Arima model

```{r}
observeEvent(input$button1,{
  req(is.null(value$series)==FALSE)
  mData$Tseries <- ts(value$series)
  
  if(any(input$checkGroup %>% grepl('^1$',.))) mData$Tseries <- tsclean(mData$Tseries)
  if(any(input$checkGroup %>% grepl('^2$',.))) mData$Tseries <- round(mData$Tseries,2)
  if(any(input$checkGroup %>% grepl('^3$',.))) mData$Tseries <- log(mData$Tseries)
  
  mData$arima<-auto.arima(mData$Tseries, ic="aic", trace=TRUE)
})
```


```{r}
output$EstimatedArima<- renderPrint({mData$arima})
verbatimTextOutput("EstimatedArima")
```


### Stationarity and differencing of time series data set:

Check of stationarity: <br>
Ho: series is non-stationary (p-value > 0.05) <br>
H1: series is stationary     (p-value < 0.05) <br> 


```{r}
#calculations block !!!!!

observeEvent(mData$arima,{

#Getting the orders of arima model
mData$p <- mData$arima$arma[1]
mData$q <- mData$arima$arma[2]
mData$d <- mData$arima$arma[length(mData$arima$arma)-1]

if(mData$d ==0){
  mData$adf<-adf.test(mData$Tseries, alternative = "stationary")
}
else {

  i=0
  for(i in mData$d-1){
  mData$Tseries<-diff(mData$Tseries, differences = 1)
  mData$adf<-adf.test(mData$Tseries, alternative = "stationary")
  }
}
mData$diffSeries<-mData$Tseries
})
```


```{r}
output$AdfTest<- renderPrint({ mData$adf })
#output$Tseries<- renderPrint({ mData$Tseries }) #nigde net
```


```{r}
verbatimTextOutput("AdfTest")
```

### Series after differencing
```{r}
output$diffplot <-renderPlotly({
  req(is.null(mData$diffSeries)==FALSE)
  diff<-ggplot()+aes_string(x=1:length(mData$diffSeries), y= mData$diffSeries) + xlab("diff series") + geom_line()
  plotly::ggplotly(diff)
}) 
```

```{r}
plotlyOutput("diffplot", height="100%")
```

### Residuals plot: ACF & PACF 
Observing residual plot and it`s ACF & PACF diagram
If ACF & PACF of the model residuals show no significant lags, the selected model is appropriate.

### ACF plot
```{r}
#plotOutput(outputId = "ACF")

#output$ACF<- renderPlot({
  #req(is.null(mData$diffSeries)==FALSE)
  #acf(mData$diffSeries)
#})

```


### PACF plot

```{r}
#plotOutput(outputId = "PACF")

#output$PACF<- renderPlot({
 # req(is.null(mData$diffSeries)==FALSE)
 # pacf(mData$diffSeries)
#})
```


### Ljung-Box test
Ljung-Box is a test of autocorrelation in which it verifies whether the autocorrelations of a time series are different from 0.<br>

More formally, the Ljung-Box test can be defined as follows:<br>
H0:	The data are random. - independent and uncorrelated;<br>
H1:	The data are not random. - remains serial correlation;<br>

Note: If the p-value is greater than 0.05 then the residuals are independent which we want for the model to be correct.<br>

```{r}
output$LBtest <- renderPrint({ 
  req(is.null(mData$arima)==FALSE)
  Box.test(residuals(mData$arima),lag=12, type="Ljung-Box") 
})
verbatimTextOutput("LBtest")
```

### Diagnostic checking
1.1) Observing residual plot and it`s ACF & PACF diagram <br>
If ACF & PACF of the model residuals show no significant lags, the selected model is appropriate.<br>
```{r}
plotOutput(outputId = "ModelResiduals")

output$ModelResiduals<- renderPlot({
  req(is.null(mData$arima)==FALSE)
  tsdisplay(residuals(mData$arima), lag.max=15, main='Model Residuals')
})
```

### ARIMA forecast

```{r}
observeEvent(mData$arima,{

  ArimaPred<-as.data.frame(forecast(mData$arima, h=4))

  if (any(input$checkGroup %>% grepl('^3$',.))){
    mData$ArimaPred<-exp(ArimaPred)
  } else {
    mData$ArimaPred<-ArimaPred
  }
})

```


```{r}
output$ArimaPrediction <- renderDataTable(

  datatable(mData$ArimaPred, filter="top")
)
```

```{r}
dataTableOutput('ArimaPrediction')
```


### Remark
Although ACF and PACF of residuals have no significant lags, the time series plot of residuals shows some cluster of volatility.<br>

It is important to note that ARIMA is a method to @linear model the data and the forecast width remains constant because the model does not reflect recent changes or incorporate new information.<br>

In other words, it provides best linear forecasts for the series, and thus plays little role in forecasting model nonlinearly.<br>

In order to model volatility, ARCH/GARCH method is used.<br>


### Testing for ARCH effect
1) Firstly, check if residual plot displays any cluster of volatility. <br>
2) Observe the squared residual plot; Are there any volatility clusters? <br>
3) Observe ACF & PACF of squared residuals. <br>

```{r}
plotOutput(outputId = "ArchEffect")

output$ArchEffect<- renderPlot({
  req(is.null(mData$arima)==FALSE)
  tsdisplay(residuals(mData$arima)^2, lag.max=15, main='Model Residuals')
})
```

### Remark
A strict white noise cannot be predicted either linearly or nonlinearly while general white noise might not be predicted linearly yet done so nonlinearly. If the residuals are strict white noise, they are independent with zero mean, normally distributed, and ACF & PACF of squared residuals displays no significant lags.


### LM test
As an alternative to Engle`s ARCH test, you can check for serial dependence (ARCH effects) in a residual series by conducting a Ljung-Box Q-test on the first m lags of the squared residual series, where m = P + Q, ARCH(P*Q), GARCH(P,Q).
```{r}
observe({
  req(is.null(mData$arima)==FALSE)
  
  LMtestTable <- data.frame(matrix(ncol = 5, nrow = 1))
  for(i in 1:5){
    u <-Box.test(residuals(mData$arima)^2,lag=i, type="Ljung-Box")
    LMtestTable[1,i]<-u$p.value
  }
  
  colnames(LMtestTable) <- c("lag 1","lag 2","lag 3","lag 4", "lag 5")
  
  mData$LMtestTable <- LMtestTable
})
```

```{r}
output$TestResultTable <- renderDataTable(
  datatable(mData$LMtestTable)
)
```

```{r}
dataTableOutput('TestResultTable')
```



### ARCH notes
The general rules for ARCH varience cofficients are following:
ht = a0 + a1 e^2 + a2 e^2 .., a0>0, 0< (sum of ai) <=1

### ARCH model selection


```{r}
observe({
req(is.null(mData$LMtestTable)==FALSE)
  
list1<-list()
list2<-list()
  for(i in 1:4){
    t<-garchFit(substitute(~arma(p,q)+garch(i,0),list(p=mData$p, q=mData$q, i=i)), data=mData$Tseries)
    t1<-attributes(t)$fit$ics
    list1[[i]]<-t
    list2[[i]]<-t1
  }

  mData$ArchModels<-list1
  #mData$crARCH<-list2
  ctr<-as.data.frame(list2)
  ctr<-as.data.frame(t(ctr))
  rownames(ctr)<-c("ARIMA-Arch(1)", "ARIMA-Arch(2)", "ARIMA-Arch(3)", "ARIMA-Arch(4)")
  mData$crARCH<-ctr
})
```


```{r}
output$ArchModelslist<- renderPrint({ mData$ArchModels
})
verbatimTextOutput("ArchModelslist")
```


### ARCH models estimated criteria
```{r}
output$ARCH_CR <- renderDataTable(
  datatable(mData$crARCH)
)
```

```{r}
dataTableOutput('ARCH_CR')
```



### GARCH model selection

```{r}
observe({
req(is.null(mData$LMtestTable)==FALSE)

list3<-list()
list4<-list()
z=1

for(i in 1:2){
  for(j in 1:2){
    garchSpec <- ugarchspec(
               variance.model=list(model="sGARCH",
                                   garchOrder=c(i,j)),
               mean.model=list(armaOrder=c(mData$p,mData$q)), 
               distribution.model="std")
    garchFit <- ugarchfit(spec=garchSpec, data=mData$Tseries)
    
    t3<-garchFit
    t4<-infocriteria(garchFit)
    
    list3[[z]]<-t3
    list4[[z]]<-t4
    
    z=z+1
    }
}

mData$GarchModels<-list3

ctrr<-as.data.frame(list4)
colnames(ctrr)<-c("ARIMA-Garch(1,1)", "ARIMA-Garch(1,2)", "ARIMA-Garch(2,1)", "ARIMA-Garch(2,2)")
mData$crGARCH<-ctrr
  
})
```


```{r}
output$GarchModelslist<- renderPrint({ mData$GarchModels
})
verbatimTextOutput("GarchModelslist")
```

### GARCH models estimated criteria

```{r}
output$GARCH_CR <- renderDataTable(
  datatable(mData$crGARCH)
)
```

```{r}
dataTableOutput('GARCH_CR')
```





### TGARCH model selection

```{r}
observe({
req(is.null(mData$LMtestTable)==FALSE)

list3<-list()
list4<-list()
z=1

for(i in 1:2){
  for(j in 1:2){
    garchSpec <- ugarchspec(
               variance.model=list(model="fGARCH",
                                   garchOrder=c(i,j),
                                   submodel="TGARCH"),
               mean.model=list(armaOrder=c(mData$p,mData$q)), 
               distribution.model="std")
    garchFit <- ugarchfit(spec=garchSpec, data=mData$Tseries)
    
    t3<-garchFit
    t4<-infocriteria(garchFit)
    
    list3[[z]]<-t3
    list4[[z]]<-t4
    
    z=z+1
    }
}

mData$TGarchModels<-list3

ctrr<-as.data.frame(list4)
colnames(ctrr)<-c("ARIMA-TGarch(1,1)", "ARIMA-TGarch(1,2)", "ARIMA-TGarch(2,1)", "ARIMA-TGarch(2,2)")
mData$crTGARCH<-ctrr
  
})
```


```{r}
output$TGarchModelslist<- renderPrint({ mData$TGarchModels
})
verbatimTextOutput("TGarchModelslist")
```


### TGARCH models estimated criteria

```{r}
output$TGARCH_CR <- renderDataTable(
  datatable(mData$crTGARCH)
)
```

```{r}
dataTableOutput('TGARCH_CR')
```


### GARCH-in-Mean model selection

```{r}
observe({
req(is.null(mData$LMtestTable)==FALSE)

list3<-list()
list4<-list()
z=1

for(i in 1:2){
  for(j in 1:2){
    garchSpec <- ugarchspec(
               variance.model=list(model="fGARCH",
                                   garchOrder=c(i,j),
                                   submodel="APARCH"),
               mean.model=list(armaOrder=c(mData$p,mData$q),
                               include.mean=TRUE,
                               archm=TRUE,
                               archpow=2), 
               distribution.model="std")
    garchFit <- ugarchfit(spec=garchSpec, data=mData$Tseries)
    
    t3<-garchFit
    t4<-infocriteria(garchFit)
    
    list3[[z]]<-t3
    list4[[z]]<-t4
    
    z=z+1
    }
}

mData$MGarchModels<-list3

ctrr<-as.data.frame(list4)
colnames(ctrr)<-c("ARIMA-MGarch(1,1)", "ARIMA-MGarch(1,2)", "ARIMA-MGarch(2,1)", "ARIMA-MGarch(2,2)")
mData$crMGARCH<-ctrr
  
})
```

```{r}
output$MGarchModelslist<- renderPrint({ mData$MGarchModels
})
verbatimTextOutput("MGarchModelslist")
```



### GARCH-In-Mean models estimated criteria

```{r}
output$MGARCH_CR <- renderDataTable(
  datatable(mData$crMGARCH)
)
```

```{r}
dataTableOutput('MGARCH_CR')
```







# ARNN

```{r}
netData <- reactiveValues(
  Seriess=NULL,
  modelfit=NULL,
  NNF=NULL,
  fcast=NULL
)
```

```{r}
actionButton("button2", "Analyze")
```


```{r}
observeEvent(input$button2,{
  req(is.null(value$series)==FALSE)
  netData$Seriess <- ts(value$series)
  
  if(any(input$checkGroup %>% grepl('^1$',.))) netData$Seriess <- tsclean(netData$Seriess)
  if(any(input$checkGroup %>% grepl('^2$',.))) netData$Seriess <- round(netData$Seriess,2)
  if(any(input$checkGroup %>% grepl('^3$',.))) netData$Seriess <- log(netData$Seriess)
  
  netData$modelfit <- nnetar(netData$Seriess, repeats = 20, lambda=0.5)
})
```

### Foreword

Feed-forward neural networks with a single hidden layer and lagged inputs for forecasting univariate time series.The nnetar function in the forecast package for R fits a neural network model to a time series with lagged values of the time series as inputs (and possibly some other exogenous inputs). So it is a nonlinear autogressive model, and it is not possible to analytically derive prediction intervals. Therefore we use simulation.The neural networks is fit by the function: <br>

nnetar(y, p, P = 1, size, repeats = 20, xreg = NULL, lambda = NULL, model = NULL, subset = NULL, scale.inputs = TRUE, x = y, …)

```{r}
output$ResultNet<- renderPrint({ netData$modelfit
})
verbatimTextOutput("ResultNet")
```


### ARNN Forecast
```{r}
observeEvent(netData$modelfit,{
  req((is.null(netData$modelfit)==FALSE))
  
  if (any(input$checkGroup %>% grepl('^3$',.))){
    netData$NNF <- forecast(netData$modelfit, PI=TRUE, h=4)
    netData$fcast <- exp(as.data.frame(netData$NNF))
  } else {
    netData$NNF <- forecast(netData$modelfit, PI=TRUE, h=4)
    netData$fcast <- as.data.frame(netData$NNF)
  }
  
})

```



```{r}
output$ARNNForecast <- renderDataTable(
  datatable(netData$fcast)
)
```

```{r}
dataTableOutput('ARNNForecast')
```


### Forecast plot
```{r}
plotOutput(outputId = "netPlot")
```

```{r}
output$netPlot <- renderPlot({
  req(is.null(netData$NNF)==FALSE)
  forecast::autoplot(netData$NNF)
})
```

